---
title: "Spotify Analysis 05/03/21"
output: 
  html_document:
    toc: true
    toc_float: true
---

INIT STUFF
```{r, message=F , echo=F}
rm(list=ls())
library(tidyverse)
library(data.table)
library(jsonlite)
library(DT)
library(scales)
commons <- function(k, n = 20, dec=T)  k %>% table(useNA="a") %>% sort(d=dec) %>% head(n)
commons <- function(k, n = 20, dec=T)  k %>% table(useNA="a") %>% sort(d=dec) %>% head(n)
p <- function(x) print(x)#data.frame(lapply(x,substr,1,32))
```

ETL
```{r}
files = paste("MyData/endsong_",0:5,".json",sep="")
df <- data.table(do.call(rbind, lapply(files, fromJSON)))
# rename columns
setnames(df,c("master_metadata_track_name","master_metadata_album_artist_name","master_metadata_album_album_name"),c("Track","Artist","Album"),skip_absent=TRUE)
df[,c("username"):=NULL]
# create columns
df[,`:=`(ts=as.POSIXct(ts,format = "%Y-%m-%dT%H:%M:%OS",tz="UTC"),
         minutesPlayed=ms_played/1000,hoursPlayed=ms_played/60000,
         Date=as.IDate(ts),Time=as.ITime(ts))]
# rearrange columns
df <- df[!is.na(Track)] %>% select(ts,Track,Artist,Album,ms_played,spotify_track_uri,everything())
# constants
TSFIRST <- min(df$ts,na.rm=T); TSLAST <- max(df$ts,na.rm=T); LENTIME <- TSFIRST - TSLAST
DATEFIRST <- min(df$Date,na.rm=T); DATELAST <- max(df$Date,na.rm=T); NDAYS <- DATELAST - DATEFIRST

# functions
hours <- function(t,s = "ms",round = 2,sum=F) {
  round(t/3600000,round)
}
minutes <- function(t,s = "ms",round = 1,sum=F) {
  round(t/60000,round)
}
seconds <- function(t,s = "ms",round = 1) {
  round(t/1000,round)
}
# clean workspace
rm(files)
```

Dates range from `r DATEFIRST` to `r DATELAST`, `r NDAYS` days. In this time I've listened to `r round(sum(df$ms_played)/1000/60/60/24,2)` days of Spotify, roughly `r round((sum(df$ms_played)/1000/60/60/24)/*100,2)`% of my total time during that period.

EDA
```{r}

# most frequent songs
df[,.(Album=first(Album),.N),by=.(Track,Artist)][order(N,decreasing=T)] %>% p

# most listen time songs
df[,.(Album=first(Album),'Hours'=round(sum(minutesPlayed)/3600,1),.N,URI=first(spotify_track_uri)),by=.(Track,Artist)][order(Hours,decreasing=T)] %>% p
max(df$ts) %>% as.numeric
# most listen time albums
msd <- max(sd(df$ts))
df[,.(Hours=round(sum(minutesPlayed)/3600,1),SD=round(sd(ts)/msd,3),Artist=first(Artist),.N,URI=first(spotify_track_uri)),by=.(Album)][order(Hours,decreasing=T)] %>% p ; rm(msd)
#%>% head(100) %>% pull(Artist) %>% commons %>% data.table

# most listen time artists
df[,.(Hours=round(sum(minutesPlayed)/3600,1),.N,URI=first(spotify_track_uri)),by=.(Artist)][order(Hours,decreasing=T)] %>% p
```


SEARCH SPECIFICS
```{r}
df[grepl("the",Track),.(.N,Hours=round(sum(minutesPlayed)/3600,2),firstListen=as.IDate(min(ts)), lastListen=as.IDate(max(ts)),URI=first(spotify_track_uri)),by=.(Track,Artist,Album)][order(Hours,decreasing=T)] %>% p

df[grepl("Nike",Track),.(Track,playLength=round(minutesPlayed/max(minutesPlayed),2),ts,URI=first(spotify_track_uri))][order(ts,decreasing=F)] %>% p

df[grepl("Blonde",Album),.(.N,Hours=round(sum(minutesPlayed)/3600,2),URI=first(spotify_track_uri)),by=.(Track,Artist,Album)] %>% p

# artist search
df[grepl("BROCKHAMPTON",Artist),.(.N,Hours=round(sum(minutesPlayed)/3600,2),URI=first(spotify_track_uri)),by=.(Track,Artist,Album)][order(Hours,decreasing=T)] %>% p
# artist search, grouped by album
df[grepl("BROCKHAMPTON",Artist),.(.N,Hours=round(sum(minutesPlayed)/3600,2),URI=first(spotify_track_uri)),by=.(Album,Artist)][order(Hours,decreasing=T)] %>% p
```


DATE & TIME
```{r}
df %>% names
ggplot() + geom_density(df,mapping = aes(x=ts))
df[,.(.N,Hours=round(sum(minutesPlayed)/3600,2),Date = first(ts)),by=.(year(Date),month(Date))]
ggplot(df[,.(.N,Hours=round(sum(minutesPlayed)/3600,2),Date = first(ts)),by=.(year(Date),month(Date))]) + 
  geom_bar(mapping = aes(x=Date)) + scale_x_date(date_labels = "%b-%d-%Y")
unique()

length(unique(format(df$ts,"%m/%y")))

# hours of music each month since 2016
df %>% { 
  ggplot(.) + 
  geom_histogram(mapping = aes(x=Date,weight=hours(ms_played)),bins=length(unique(format(.$ts,"%m/%y")))) + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "3 month") + 
  ylab("Hours of Music") + xlab("Month") +
  theme(axis.text.x = element_text(angle=45,hjust=1))
}

# mean hours listened 
df[order(ts),.(Hours=hours(sum(ms_played))),by=.(Month = format(ts,"%m/%y"))][,mean(Hours),by=.(Month = substr(Month,1,2))][order(as.numeric(Month))]

# %>% ggplot(.) + geom_bar(aes(x=Month,weight=Hours)) + 
  #scale_x_date(date_labels = "%b-%y", date_breaks = "3 month") + 
  ylab("Hours of Music") + xlab("Month") +
  theme(axis.text.x = element_text(angle=45,hjust=1))

# What days did I listen to music? What percent of days?
merge(data.table(Date=as.IDate(DATEFIRST:DATELAST)),df[,.(minutesPlayed = sum(ms_played),!.N==0),by=.(Date)],on=.("Date"))
df[,.(Hours=sum(minutesPlayed)/3600),by=.(Date)] %>% nrow
1686
(1546/1686*100) %>% round(2) 
```


```{r}
# # time of day
# 
# ggplot() + geom_density(df,mapping = aes(x=as.ITime(ts))) +
#   scale_x_time(name = "Time of Day", time_labels = "%h:%m") +
#   ylab("Density")
```

Cluster songs, where distance = their temporal distance


- Blonde
- Worlds
- Homey
- Modal Soul
- Kid A
- good kid, m.A.A.d city
- The Human Condition
- Spiritual State
- To Pimp A Butterfly
- Metaphorical Music
- Tapestry

